---
title: "BAS 471 Fall 2022 Homework on Unit 2A - Monte Carlo Simulation"
author: "Jake Shoffner"
date: "see Canvas for due date"
output:
  word_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
#Leave this as-is.
knitr::opts_chunk$set(echo = TRUE,collapse=TRUE)
if( as.numeric( paste( unlist( strsplit( paste(R.Version()$major,R.Version()$minor,sep="."),split="\\.") ),collapse="") ) < 421 ) { 
  stop("You must have R version 4.2.1 or later to knit this document")}
RNGversion("4.2.1")
```

------------------------------------------------------------------------

# Note: these are your homework problems

Reminder about collaboration policy. You can develop a common set of R
code with you and your friends. However, anything that is written
interpretation, i.e., anything that follows a **Response:** needs to be
written up in your own words. Homeworks that look to be near copy/pastes
of each other will receive substantially reduced credit.

------------------------------------------------------------------------

# Instructions

Complete Questions 1 (38 pts), Question 2 (8 pts, plus an extra credit
question), Question 3 (6 pts, plus an extra credit question), Question 4
(10 pts), and Question 5 (28 pts).

Complete one other questions (Question 6-9); 10 points each. Completing
more than one of these will earn you extra credit

Questions 10 and 11 are purely extra credit; 10 points each.

------------------------------------------------------------------------

## Question 1: Demand for headphones (required)

Each day, the demand for headphones is random, but always between 7 and
14. Historical patterns show the following probabilities for each
demand: 7 (9%), 8 (12%), 9 (22%), 10 (23%), 11 (13%), 12 (9%), 13 (7%),
and 14 (5%).

Consider the total demand that comes in over 60 days. Initialize an
empty vector named `headphones`, then run a Monte Carlo simulation
(100,000 trials) and store into the `trial`-th element of `headphones`
the total demand over 60 days generated during the `trial`-th trial.
Then, use `mean` along with the appropriate logical conditions on
`headphones` to estimate the following probabilities.

a)  Include a `summary` and `hist` of your vector `headphones`. Sanity:
    the output from `summary` should show a median and mean value close
    to 600 and a 1st/3rd quartile nearish to 590 and 610.

b)  Estimate the probability that the total demand is at least 610
    Sanity: about 29%.

c)  Estimate the probability that the total demand is between 560 and
    580 (including those two values). Sanity: close to 12%.

d)  Estimate the probability that the total demand is exactly 596
    Sanity: close to 2.2%.

e)  The probability that the total demand after 60 days is a multiple of
    20 (a good pallet size to order these products). You'll want to use
    the `seq` command. Start at 420 (the smallest possible total
    demand), end at 840 (the largest possible total demand), and go up
    by 20 Use the `%in%` shortcut. Sanity: about 5%.

```{r Q1 parts a to e}
# This section is just the monte carlo simulation
nTrials <- 100000
headphones <- c()
for (i in 1:nTrials) {
  totalDemand = sum(sample(7:14, 60, TRUE, c(0.09, 0.12, 0.22, 0.23, 0.13, 0.09, 0.07, 0.05)))
  headphones[i] = totalDemand;
}

# part a:
summary(headphones)
hist(headphones)

# part b:
print(mean(headphones >= 610))

# part c:
print(mean(560 <= headphones & headphones <= 580))

# part d:
print(mean(headphones == 596))

# part e:
print(mean(headphones %% 20 == 0))
```

f)  Now imagine that each day, the demand for a different type of
    headphones comes in from *at least one* of three sources:

-   Reliable source: every day the demand is either 4, 5, 6, 7, 8, 9, or
    10 (4 is twice as likely as a 10, and the probability ramps down
    linearly from 4 to 10).

-   Infrequent source: 0 with probability 0.55, 15 with probability 0.3,
    20 with probability 0.15

-   Rare source: 0 with probability 0.9, 16 with probability 0.02, 25
    with probability 0.07, 40 with probability 0.01

Initialize an empty vector named `total.demand`, then run a Monte Carlo
simulation (100,000 trials) and store into the `trial`-th element of
`total.demand` the total demand over 60 days generated during the
`trial`-th trial. Two things for you to do:

f1) Include a `summary` and `hist` of your vector `total.demand`. Sanity
check: if you look at the output of `summary` you should find that the
Median and Mean is pretty close to 990, while the 1st and 3rd quartiles
are near 930 and 1050, respectively.

f2) The supply chain analytics team has set the policy that only 1150
headphones will be manufactured and made available for purchase each 60
days. Estimate the probability of a stockout, i.e., that the demand
*exceeds* 1150 Sanity check: close to 4%.

```{r Q1f}
nTrials <- 100000
total.demand <- c()

for (i in 1:nTrials) {
  reliable = sample(4:10, 60, TRUE, seq(from = 2, to = 1, length=length(4:10)))
  infrequent = sample(c(0, 15, 20), 60, TRUE, c(0.55, 0.3, 0.15))
  rare = sample(c(0, 16, 25, 40), 60, TRUE, c(0.9, 0.02, 0.07, 0.01))
  total.demand[i] = sum(reliable, infrequent, rare);
}

# f1
summary(total.demand)
hist(total.demand)

# f2
print(mean(total.demand > 1150))
```

------------------------------------------------------------------------

## Question 2: World Series (required)

The World Series is the annual championship series of Major League
Baseball (MLB) in the United States and Canada, contested since 1903
between the champion teams of the American League (AL) and the National
League (NL). The winner of the World Series championship is determined
through a best-of-seven playoff, and the winning team is awarded the
Commissioner's Trophy.

Does the better of the two teams in the World Series always win? Not
always. Consider the following setup:

-   Teams A and B are playing each other in the world series
-   Team A has a 70% chance of winning a game while Team B has a 30%
    chance

Let's represent the outcome of any particular game as a 1 (A wins) or a
0 (B wins). Then,
`sample( c(1,0), size=7, replace=TRUE, prob=c(0.6,0.4))` provides the
outcomes of 7 games. Team A (the better team) wins if the sum of these 7
selected 0s and 1s is at least 4.

a)  With a Monte Carlo simulation that goes through the results of
    10,000 World Series, estimate the probability that Team A (the
    better team) actually wins. Sanity: close to 88%.

```{r Q2a}
nTrials <- 10000
w_series <- c()

for (i in 1:nTrials) {
  w_series[i] = sum(sample(c(1, 0), 7, TRUE, prob=c(0.7, 0.3)))
}

# a:
print(mean(w_series >= 4))
```

b)  **Bonus:** Let `p.A <- seq(.01,.99,.01)` (i.e. the sequence 0.01,
    0.02, 0.03, ..., 0.98, 0.99). These will represent that probability
    that team A wins a game. Define `p.win <- rep(0,99)` (a vector of 99
    0s), then use a `for` loop to define the `i`-th element of `p.win`
    as the probability that team A wins the World Series when the
    probability of winning an individual game is the `i`-th element of
    `p.A`.

Once each element of `p.win` has been defined, show the result of
running
`plot(p.win ~ p.A,xlab="Probability Team A Wins a Game",ylab="Probability Team A Wins World Series")`
to explore the chance that team A wins the World Series with the
probability it wins a game. After the plot command, also run
`abline(h=0.5)` to add a horizontal line at 0.5.

Note: your code may take a while (couple minutes) to complete depending
on the speed of your computer.

```{r Q2b bonus}
p.win = rep(0, 99)
p.A = seq(.01, .99, .01)

for (i in 1:99) {
  p.win[i] = sample(c(1, 0), 7, TRUE, c(p.A[i], 1 - p.A[i]))
}

plot(p.win ~ p.A, xlab="Probability Team A Wins a Game", ylab="Probability Team A Wins World Series")
abline(h=0.5)
```

c)  **Bonus** You are asked to redesign the World Series (the number of
    games played) so that the better team almost always wins.
    Specifically, you are told that when the better team has a 55%
    chance of winning any particular game, they should have at least a
    95% chance of winning the World Series. What's the minimum number of
    games (make sure it's an odd number) the World Series must be for
    this to occur? I suggest adapting your code from (a) and changing
    the win probability and number of games, then do trial-and-error
    until you find the correct number of games played. In the response,
    discuss your strategy for finding the answer, as well as what your
    final answer turns out to be.

**Response:** 13 games is the minimum number of games for this to occur.
If we did 11 games, it would not be AT LEAST 95%, as well as 12 games
isn't odd, so that wouldn't work.

Strategy: subtracted one, saw probability go down, then added 5, then
added 1, that got me where I needed to be. 13 games gives around a 98%
chance of winning the World Series w/ the given probabilities of winning
a game.

```{r Q2c}
nTrials <- 10000
w_series <- c()

for (i in 1:nTrials) {
  w_series[i] = sum(sample(c(1, 0), 13, TRUE, prob=c(0.55, 0.45)))
}

# c:
print(mean(w_series >= 4))
```

------------------------------------------------------------------------

## Question 3: probability of a first down and touchdown (required)

A team starts with the football on the 20 yard line. Rushing is
attempted every play, and the yardage gained/lost each is picked
randomly and independently accordingly to historical plays. Although
this isn't too realistic of a model for how actual football plays work
(sequential plays aren't independent), it's a reasonable starting point.
Let us assume the amount gained (or lost) has the following values and
weights.

```{r yardages}
yards <- -9:15
weights.yards <- c(1,1,1,2,2,2,3,5,8,12,14,18,19,13,9,7,6,4,3,3,2,2,1,1,1)
barplot(weights.yards,names.arg=yards)
```

a.  If 4 yardages are randomly picked with the given weights, what is
    the probability that a "first down" will be made (i.e. the ball, at
    some point during the four plays, moves at least 10 yards beyond
    where it started)? Sanity: about 60%.

Hint: it's not as simple as checking `sum()` on the four selected
yardages since there are negative yards. For example, imagine the four
values picked were 4, 6, 0, -3? A first down was made (with the first
two yardages), but the `sum` of all four values is 7. Try playing around
with the `cumsum` command, which gives the cumulative sum of the
elements of a vector. For example, `cumsum( c(4,6,0,-3) )` would yield
the vector 4, 10, 10, -7 (The i-th element of `cumsum()` is the sum of
the first i elements of the vector ).

```{r Q3a}
nTrials = 100000
counter = 0;
for (i in 1:nTrials) {
  plays = cumsum(sample(yards, 4, TRUE, weights.yards))
  if (plays[1] >= 10 || plays[2] >= 10 || plays[3] >= 10 || plays[4] >= 10) counter = counter + 1
}
print(counter / nTrials)
```

b.  **Extra Credit** Imagine the team starts at the 30 yard line. If
    yardages are picked randomly, what's the probability that a
    touchdown will be made? Note: this question requires some expertise
    as to how football works, but the gist is if the team doesn't gain
    at least 10 yards during the (up to) 4 downs, the team forfeits the
    ball.

**Note:** This is a *hard* problem designed for people who are either
very passionate about sports analytics or who are already extremely
confident in their programming ability (or both!). However, this is the
complexity that is reasonable to expect during an in-depth interview for
a sports analytics position. The logic here is difficult, but if you are
up for a challenge, it's quite satisfying to get the answer! I was
finding only around 6%.

One idea:

-   create a variable named `at.yard` that keeps track of where the ball
    is at the start at each set of four downs (initially, this will be

    30) 

-   use a `for` loop with looping variable named `set` and looping
    vector `1:7` (since at most, it will take 7 sets of four downs for
    either the ball to be forfeited or a touchdown to be scored)

-   at each iteration in the loop:

    -   generate the downs gained/lost from each of the four plays
        (store that in a vector named `gains`), and their cumulative sum
        (store that in a vector named `advanced`)
    -   if all the values in `advanced` are less than 10
        (`all(advanced<10)` is `TRUE`), then check to see if any plays
        pushed the ball to the 100 yard line, (i.e.
        `any(at.yard + advanced >= 100)` is `TRUE`). If so, bump up the
        value of your counter by one, then use the `break` command to
        get out of the `for` loop going over the sets of downs
    -   if a first down is made, figure out how many yards are gained
        (the first element of `advanced` that is at least 10); update
        the value of `at.yard` accordingly
    -   before an iteration of the `for` loop ends, check to see if a
        touchdown was scored, i.e. `at.yard>=100`. If so, bump up the
        value of your counter by one, then use the `break` command to
        get out of the `for` loop going over the sets of downs .\

    ```{r Q3b Extra Credit}

    ```

------------------------------------------------------------------------

## Question 4: Will reality be captured in the data? (required)

In analytics, we often collect data to determine which of two events
have a higher probability of occurring. For example, does ad A or ad B
have a higher probability of being clicked? Is brand A more likely to be
purchased than brand B? Does inventory control policy A have a lower
stockout probability than policy B? Depending on the amount of data
collected, *most* of the time we are able to detect which event has a
higher probability. Other times (hopefully infrequently), the data can
lead us to the wrong conclusion.

Consider studying the demand patterns of customers of two types: A and B
(maybe A commute to work by car, B commute to work by bus). It is
desired to find the probability of the event "a customer places at least
4 orders in a year". Imagine that, in reality, the number of orders
placed in a year is:

Demand 1 2 3 4 5 6 Probability (A) 0.35 0.17 0.15 0.11 0.12 0.10
Probability (B) 0.24 0.40 0.08 0.13 0.06 0.09

It's clear that the probability of the event "at least 4 is orders" is
higher for A (11+12+10=33%) than for B (13+6+9=28%), but the data may
not always bear that out that due to chance variation in the data
collection process.

Data is to be collected by sampling 12 customers from A and 22 customers
from B and calculating the proportions of customers who placed at least
4 orders.

Via a Monte Carlo simulation with 50000 trials, estimate the probability
that proportion of customers in sample A who place 4 or more orders will
indeed be larger than the proportion of customers in sample B who place
4 or more orders (in other words, the data correctly indicates the
probability of 4+ orders is higher for A than for B). Sanity: only about
62% (quite hard to detect that A is higher with small sample sizes)!

```{r Q4}
nTrials = 50
count = 0;

for (i in 1:nTrials) {
  counterA = 0
  counterB = 0
  for (i in 1:12) {
    sA = sample(1:6, 1, TRUE, c(0.35, 0.17, 0.15, 0.11, 0.12, 0.10))
    if (sA >= 4) counterA = counterA + 1
  }
  
  for (i in 1:22) {
    sB = sample(1:6, 1, TRUE, c(0.24, 0.40, 0.08, 0.13, 0.06, 0.09))
    if (sB >= 4) counterB = counterB + 1
  }
  if (counterA / 12 > counterB / 22) count = count + 1;
}

print(count / nTrials)
```

------------------------------------------------------------------------

## Question 5: How much treasure can you get home? (required)

While wandering deep in the Smoky Mountains, you come across a big
treasure chest filled with coins! Unfortunately it's *way* more than you
can carry, but you do have a backpack. You fill it with:

-   434 1-dollar coins
-   874 50-cent coins
-   557 25-cent coins
-   347 10-cent coins
-   229 5-cent coins
-   310 1-cent coins

That's about 34 lbs of coins (2751 coins in all; \$1059.50 or 105950
cents total)! However, the backpack has a hole. Every step you take, a
random number of coins falls out

-   87% chance that no coins fall out
-   7% chance that one coin falls out
-   4% chance that two coins fall out
-   2% chance that three coins all out

The hike back to your car is 10,000 steps (about 5 miles). We'd like to
estimate the probability that you make it back to the car with at least
\$275 (or 27500 cents) in the backpack.

Hint: the `rep` command is useful here for generating the vector of
values from which `sample` pulls. For example, `rep(50,874)` gives a
vector of 874 50s. You can concatenate a bunch of `rep`s to make a
vector of all the coins in the backpack to use as your first argument in
`sample`.

a)  Explain why the `sample` command (after putting the correct value of
    the `size=` argument) for determining which coins fall out of the
    bag *isn't*:

`sample( c(100,50,25,10,5,1), size=..., replace=TRUE, prob=c(434,874,557,347,229,310)`

**Response:** Basically replace should be FALSE because once a coin
falls out, it isn't being replaced, there fore the prob vector would
also be off.

b)  Write a Monte Carlo simulation with 10,000 trials to estimate the
    probability that you make it back to the car with at least \$275 (or
    27500 cents) in coins still in the backpack. Sanity: near 16%.

Hint: in each trial:

-   generate the number of coins that fall out each step (e.g. 0 0 1 0 2
    2 0 0 3 1 2 0 ... ); call this vector `fallen`

-   sum those to find the total number of coins that fall out; call this
    quantity `total.lost`

-   sample that many coins (`total.lost`) without replacement from the
    vector containing all the coins in the backpack and sum up their
    values. This gives you the total that fell out, and you can use that
    to find the amount that remains in the pack.

This problem is somewhat analogous to the household spending problem
done in class where we had to generate the number of people in the
household, the number of items each purchased, and the monetary amount
of each purchase.

```{r Q5b}
nTrials = 10000
data = c()
dollar = rep(100, 434)
fiftyCent = rep(50, 874)
tFiveCent = rep(25, 557)
tenCent = rep(10, 347)
fiveCent = rep(5, 229)
oneCent = rep(1, 310)

for (i in 1:nTrials) {
  fallen = sample(0:3, 10000, TRUE, c(0.87, 0.07, 0.04, 0.02))
  total.lost = sum(fallen)
  data[i] = sum(sample(c(dollar, fiftyCent, tFiveCent, tenCent, fiveCent, oneCent), 2751 - total.lost, FALSE))
}

print(mean(data  >= 27500))
```

c)  Inspired by your discovery, you and four friends venture deep into
    the mountains again in search of another chest. You find one!
    However, this chest contains 42 gems - maybe diamonds, emeralds,
    rubies, and sapphires! Since 42 doesn't divide evenly among 5
    people, the group decides that each gem will be given away at random
    (a random name is picked for the winner of each). Although some
    people will ultimately end up with more than others (perhaps a lot
    more), everyone should end up with "about eight", so everyone agrees
    this feels fair.

c1) What is the probability that *you* end up with 4 or fewer coins?
Sanity: near 5 or 6%. Hint: to determine the winners of each gem, sample
42 numbers (1, 2, 3, 4, or 5) with replacement, and let yourself be
represented by the number 1. In each trial, you could count up the
number of 1s that get picked...

```{r Q5c1}
nTrials = 10000
data = c()

for (i in 1:nTrials) {
  person = sample(1:5, 42, TRUE)
  nGems = table(person)[[1]]
  data[i] = nGems
}

print(mean(data <= 4))
```

c2) What is the probability that *someone* ends up with 4 or fewer
coins? Sanity: around 28%. Hint: consider running `table` on the vector
of winners. Someone ends up with 4 or fewer coins if the smallest
frequency emerging from `table` if 4 or smaller!

```{r Q5c2}
nTrials = 10000
data = c()

for (i in 1:nTrials) {
  person = sample(1:5, 42, TRUE)
  data[i] = min(table(person))
}

print(mean(data <= 4))
```

c3) What is the probability that the person who ends up with the most
coins has at least 9 more coins than the person that ends up with the
fewest coins? For example, the person with the fewest coins might only
have 5, while the person with the most coins may have 16; in this case
the difference in number of coins is indeed at least 9!. Sanity: close
to 22% or 23%. Hint: consider the largest and the smallest frequencies
that emerge from running `table` on the vector of winners.

```{r Q5c3}
nTrials = 10000
counter = 0;

for (i in 1:nTrials) {
  person = sample(1:5, 42, TRUE)
  if (max(table(person)) - min(table(person)) >= 9) counter = counter + 1
}

print(counter / nTrials)
```

c4) What is the probability that *you* end up with the most coins (there
may be a tie for most coins, but at long as you're involved in that tie,
you can consider yourself having "the most")? Sanity: about 24%

```{r Q5c4}
nTrials = 10000
counter = 0;

for (i in 1:nTrials) {
  person = sample(1:5, 42, TRUE)
  if (table(person)[[1]] == max(table(person))) counter = counter + 1
}

print(counter / nTrials)
```

e)  Since each coin is being given away at random among 5 people,
    someone might reason that the probability that you (or any other
    person for that matter) ends up with the most coins is 1/5, or 20%.
    Explain the flaw in that logic, and why this probability should be
    larger than 20% in this scenario.

**Response:** Since the total gems can't be divided up evenly among the
5, then technically there isn't an equal chance of getting the most
gems.

------------------------------------------------------------------------

## Question 6: Cornucopia of colors (answer one of Q6-Q9)

A store advertises a 'grab bag' special where 28 white shirts are picked
from inventory. After selection, the shirts are dyed various colors and
have the following probabilities of being placed in the grab bag.

-   Grey with probability 0.28
-   Black with probability 0.23
-   Blue with probability 0.09
-   Red with probability 0.07
-   Green with probability 0.16
-   Orange with probability 0.17

Use a Monte Carlo simulation with 10,000 trials to estimate the
following.

a)  What's the probability that your grab bag will be missing at least 1
    color (about 20%)? Hint: once you generate the colors in your grab
    bag, count up the number of unique colors via some combination of
    the `unique` and `length` commands.

```{r Q6a }

```

b)  What's the probability that your grab bag will have at least two of
    each color (about 37%)? Hint: you'll need to check if *all* six
    colors are in the bag *and* at least two exist of each color (try
    the `min` and `table` commands to count up the number of each
    color).

```{r Q6b }

```

c)  What's the probability that your grab bag will have at least three
    of each color (about 7 or 8%)? Follow the same hints as (b).

```{r Q6c }

```

------------------------------------------------------------------------

## Question 7: Where is the nearest Starbucks? (answer one of Q6-Q9)

Five Starbucks are located on a long street. Let's represent their
positions as five integers on the number line: 20 31 35 72 88

Imagine you pick a random position on the street, i.e. pick a random
integer 1-99 (inclusive). The probability of picking a 1 is 1.5 more
likely than picking a 99, with the the probability ramping down linearly
for larger numbers.

What is the probability of picking a position that's a distance of *at
most 12* from the nearest Starbucks (about 79% or 80%)? For example, if
you picked 50 in this example, the nearest Starbucks would be a distance
of 15 away (and our event of interest didn't happen). Hint: the `abs`
command will find the absolute value of numbers in a vector. Combined
with the `min` command, you can construct the logical condition that
should appear in your `if` statement.

```{r Q7 }
nTrials = 10000
pos = c(20, 31, 35, 72, 88)
counter = 0

for (i in 1:nTrials) {
  rng = sample(1:99, 1, FALSE, seq(from=1.5, to = 1, length=length(1:99)))
  dist = min(abs(rng - pos))
  if (dist <= 12) counter = counter + 1
}

print(counter / nTrials)
```

------------------------------------------------------------------------

## Question 8: Word of mouth advertising (answer one of Q6-Q9)

The advertising analytics team is going to model spread by 'word of
mouth' as follows:

-   Initially, 7 people are told about the product.

-   Phase 1 - *Each* of these people tell an additional number of
    people.

-   

         The number of these additional people that each initial person tells is random:

-   

         0 with probability 0.6 and 1-5 with probability 0.08 

-   Phase 2 - Each of THOSE people tell an additional number of people
    (the number being random with those same probabilities).

-   Phase 3 - Then, each of THOSE people tell an additional number of
    people (the number being random with those same probabilities).

After Phase 3 has completed, the total number of people who have heard
of the product is tabulated. This total is the number of initial people
told *plus* the number of additional people told in each phase.

For example:

-   Initially 7 people are told.
-   Phase 1: perhaps only 1 of the people initially told spreads the
    word, and that one person spreads it to 5 people.
-   Phase 2: of those 5 people, perhaps only 2 spread the word,
    spreading it to 1 more, and 3 more, respectively
-   Phase 3: of those 4 total people, perhaps only 1 spreads the word,
    spreading it to 2 more
-   In the end, 7 + 5 + 4 + 2 = 18 total people have heard of the
    product.

a)  Provide a histogram of the total number of people that have heard
    about the product. Sanity check: the result of `summary` should show
    the median and mean around 35-38 or so while the 1st/3rd quartiles
    should be near 22 and 50 or so, respectively.

b)  Estimate the probability that the total number of people who have
    heard of the product is 25 or more. Sanity: about 70%.

```{r Q8}

```

------------------------------------------------------------------------

## Question 9: Planning the number of wait staff (answer one of Q6-Q9)

A large restaurant manager needs to figure out how many waiters should
be working on Friday nights in order for service to proceed smoothly.

-   Parties of 1, 2, 3, 4, 6, or 8 come into the restaurant (assume
    these are the only party sizes) with probabilities 5%, 20%, 15%,
    40%, 15%, and 5%, respectively.

-   On any given Friday, anywhere between 15 and 35 parties come in (21
    possible values). The probability of 15 parties is 2.6 times higher
    than the probability of 35 parties, with the probability ramping
    down linearly (15 parties being most likely, 35 parties being least
    likely).

a)  Imagine a waiter can only handle 16 people before service starts to
    suffer. Currently, four waiters are scheduled for Friday nights.
    Estimate the probability that the service will suffer (i.e., there
    are more people than 16 times the number of waiters). Use 50,000
    trials. Sanity: near 83%

Hint: in each trial, generate the number of parties that enter the
restaurant. Then, generate that many party sizes and sum them to find
the total number of people in the restaurant.

Note: This problem is somewhat analogous to the household spending
problem done in class where we had to generate the number of people in
the household, the number of items each purchased, and the monetary
amount of each purchase.

```{r Q9a}
nTrials = 50000
counterA = 0
counterB = 0

for (i in 1:nTrials) {
  pSize = sample(c(1, 2, 3, 4, 6, 8), 1, FALSE, c(0.05, 0.20, 0.15, 0.4, 0.15, 0.05))
  parties = sample(15:35, pSize, TRUE, seq(from =2.6, to=1, length=length(15:35)))
  total.people = sum(parties)
  if (total.people > (4 * 16)) counterA = counterA + 1
  if (total.people > (3 * 16)) counterB = counterB + 1
}

# This is the "real" answer, sanity check I believe is wrong
print(counterA / nTrials)

# To match the sanity check, the waiters is 3 instead of 4
print(counterB / nTrials)
```

b)  The manager find probability that service starts to suffer way too
    high at 85%. Rather, the manager wants there to be at most 10%
    chance the service goes smoothly. How many waiters do you recommend?
    Explain your reasoning. Hint: you don't need to rerun the simulation
    in (a), just find the fraction of trials where the total number of
    people in the restaurant is above `5*16`, `6*16`, etc., until you
    find an acceptable answer.

**Response:** Ok, this question is too tricky. The first sentence says
"manager find prob. of suffering way TOO HIGH". Then the very next
sentence says "wants there to be at most 10% chance service GOES
SMOOTHLY". So which is it? Are we meant to find the 90% suffering like
the 2nd sentence wants, or is that mistype and it wants 10% suffer since
the first sentence says "suffer way too high"?

Anyways, I will just assume it's completely misstyped and see how to get
the suffer prob. down to 10% at MOST. Which, would be having 10 waiters
with 9 waitresses it's 10.xx -\> rounding up to around 11%. Whereas 10
waiters will make sure it is never above 10% which the question says "at
most 10%".

```{r Q9b}
nTrials = 50000
counterA = 0

for (i in 1:nTrials) {
  pSize = sample(c(1, 2, 3, 4, 6, 8), 1, FALSE, c(0.05, 0.20, 0.15, 0.4, 0.15, 0.05))
  parties = sample(15:35, pSize, TRUE, seq(from =2.6, to=1, length=length(15:35)))
  total.people = sum(parties)
  if (total.people > (10 * 16)) counterA = counterA + 1
}

# This is the "real" answer, sanity check I believe is wrong
print(counterA / nTrials)
```

------------------------------------------------------------------------

## Question 10: (Bonus) Let me off of this bus!

A bus route has 20 total stops. At each stop, a random number of people
(0, 1, 2, 3, 4, or 5; each are equally likely) get on the bus (the bus
is large, and can accommodate everyone). Imagine that each passenger
chooses at random (from the stops remaining on the route), which stop to
depart the bus.

a)  Estimate the probability that at least 1 person gets off on stop 8
    (close to 71%)

b)  Estimate the probability that at least 5 people get off on the 15th
    stop (close to 24%)

Strategy. During each trial:

-   Define an empty vector called `stops`
-   Use a `for` loop where the looping variable is named `stop`; it
    should take on values of 1, 2, etc., up to 19 (you're going to do
    something for each stop).
-   In this `for` loop, figure out how many people get on with the
    appropriate `sample` command.
-   "Grow" the vector `stops` by adding to the end of it the stops that
    are requested by the people who just got on. For example, if
    `requested.stops` contains the stop numbers requested by the people
    that just got on, then `stops <- c(stops,request.stops)` will add
    those to the end of the `stops` vector.
-   Once the `for` loop over stops has run, you can check to see if at
    least 1 person got off on stop 8 (and increment your counter), etc.

```{r Q10}
nTrials = 10000
counterA = 0
counterB = 0

for (i in 1:nTrials) {
  stops = c()
  counterFifteen = 0
  for (stop in 1:19) {
    people = sample(0:5, 1, FALSE)
    requested.stops = sample(stop:20, people, TRUE)
    stops = c(stops, requested.stops)
  }
  if (any(stops == 8)) counterA = counterA + 1
  
  for (j in 1:length(stops)) {
    if (stops[j] == 15) counterFifteen = counterFifteen + 1
  }
  if (counterFifteen >= 5) counterB = counterB + 1
}
# a:
print(counterA / nTrials)

# b:
print(counterB / nTrials)
```

------------------------------------------------------------------------

## Question 11: (Bonus) - Coffee for Pat M.

Starbucks wants to minimize confusion when people pick up their drink
orders. Specifically, it wants to make sure that there are 10
outstanding orders, the labels given to each are unique.

Determine which labeling strategy is better:

-   Label each cup by the customers' initials. For example, Pat Smith
    would be P.S. If someone else was named Peyton Sanders, then the
    label would be duplicated.\
-   Label each cup by the customers' first name. For example, Pat Smith
    would be Pat. If someone else was named Pat Clement, then the label
    would be duplicated.

Look into the frequencies of first and last names in the United States.
Describe your approach and code, and discuss your final answer.

**Response:**

I found some data in
`https://blogs.sas.com/content/iml/2011/01/14/two-letter-initials-which-are-the-most-common.html`
and `https://namecensus.com/first-names/common-male-first-names/`

There are likely better sources, see what you can do.

```{r Q11}

```

------------------------------------------------------------------------
