---
title: "BAS 471 Fall 2022 Homework on Unit 5 Parts 1-3 - Probability Models for Counts"
author: "Jake Shoffner"
date: "see Canvas for due date"
output:
  word_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
#Leave this as-is.
knitr::opts_chunk$set(echo = TRUE,collapse=TRUE)
if( as.numeric( paste( unlist( strsplit( paste(R.Version()$major,R.Version()$minor,sep="."),split="\\.") ),collapse="") ) < 421 ) { 
  stop("You must have R version 4.2.1 or later to knit this document")}
RNGversion("4.2.1")
```

------------------------------------------------------------------------

# Note: these are your homework problems

Reminder about collaboration policy. You can develop a common set of R
code with you and your friends. However, anything that is written
interpretation, i.e., anything that follows a **Response:** needs to be
written up in your own words. Homeworks that look to be near copy/pastes
of each other will receive substantially reduced credit.

------------------------------------------------------------------------

## Question 1

The following problems ask you to work with summation or product
notation. For these, I suggest writing out the equation by hand, taking
a photo, and inserting it into your Word document after you've knitted
the homework. Though not recommended, you can also write these equations
in Word using the equation editor, or learn how to do this is LaTeX.
Although it's not required, I suggest you also evaluate the sum with
Wolfram-Alpha since we will begin heavily relying on it for the
remainder of the semester.

Quick summary (though summation notation is addressed in the notes and
videos):
`https://math.illinoisstate.edu/day/courses/old/305/contentsummationnotation.html`

a)  Write the following sum using summation notation, having the
    summation index `i` start at 10. Then, in the R chunk, evaluate the
    sum by defining `i` to be the appropriate integer sequence, and by
    using `sum` on the appropriate equation using `i`. Sanity check: a
    little over 168000.

32 + 35 + 38 + 41 + 44 + .... + 998 + 1001 + 1004

**Response:** Work below in R chunk.

```{r Q1a}
sum = 0;
for (i in seq(from = 10, to = 982, by= 3)) {
  sum = sum + (i + 22)
}
sum
```

b)  Write the following infinite sum using summation notation, then use
    R to evaluate it:

$$1 - \frac{3}{5} + \frac{4}{10} - \frac{5}{17} + \frac{6}{26} - \ldots$$

Hints: note how each term looks like the value `(i+1)/(1+i^2)` for some
value of `i`. To get the alternating signs, you could multiply each term
by either $(-1)^i$ or $(-1)^(i+1)$, depending on what value of `i` you
use to start out the series. Since there is an infinite number of terms,
to "evaluate" the sum in R, define `i` to be an integer sequence with at
least 500,000 elements, then `sum` the appropriate equation using `i`.
Sanity check: a little above 0.633.

**Response:** Work below in R chunk.

```{r Q1b}
sumB = 0;
for (i in 1:500000) {
  sumB = sumB + (-1)^(i - 1) * ((i + 1) / (1 + i ^ 2))
}
sumB
```

c)  Consider the product:

$$\frac{2}{1} \cdot \frac{4}{3} \cdot \frac{6}{5} \cdot  \frac{8}{7} \ldots \cdot \frac{1000}{999}$$

Notice the numerators are multiples of 2 and the denominators are a
sequence of odd numbers. Write the equation in product notation,
evaluate it in R. Sanity check: a little under 40.

**Response:** Work below in R chunk.

```{r Q1c}
# needs to be 1 since we are doing product instead of sum
sumC = 1;
for (i in 1:999) {
  # if it's an odd number, then we can use this here
  if (i %% 2 == 1) sumC = sumC * ((i + 1) / i)
}
sumC
```

------------------------------------------------------------------------

## Question 2

When signing up as a new customer to a streaming service, you are asked
to select at least three shows from a selection of 80 different shows
hosted by the streaming service.

Let `X` be the number of different shows selected by a new customer. The
possible values for `X` are between 3 and 80. The analytics team wants
you to explore making a probability model for `X`.

The team has a shape they'd like the probabilities to follow:

$$Shape = 2e^{-(x-3)/5} + 0.01xe^{-(x-30)^2/200} \qquad x=3, 4, \ldots, 79, 80$$
This shape is inspired by the fact that upon signing up, it's the most
likely that someone selects the minimum of three shows. Then, after
three, larger values are progressively more rare until there's a "bump"
around 30 or so (this tends to be where more die-hard new customers end
up selecting a lot of shows). The following R chunk defines a vector `x`
to contain the possible values of `X`, and defines a vector `shape` that
gives the relative frequencies of the possible values in `x`. Remember
that to take `e` and raise it to a power in R, you need to write it as
`exp()`, with the power going inside the parentheses.

```{r Q2}
x <- 3:80 
shape <-  2*exp(-(x-3)/5) + 0.01*x*exp( -(x-30)^2 / 200 )
barplot(shape,names.arg=x,xlab="Total Selections",ylab="Relative Frequency")
```

a.  The values in `shape` clearly do not contain the *probabilities* of
    each possible value (among other reasons, some values in `shape` are
    greater than 1). However, since there are a finite number of
    possible values of `x`, we can convert the `shape` vector into a
    vector of probabilities using the standard conversion method! Do so,
    defining a vector `p` whose elements give the probabilities of
    observing each value of `x`. Show the sum of the elements in `p`
    equals 1, print to the screen the value of `p[30]` (it should be
    0.01723217), and provide a barplot of the PMF. Label the horizontal
    axis "Total Selections" and the vertical axis "Probability.

```{r Q2a}
# scaling it to probabilities
p = shape / sum(shape)

# sum of p = 1
sum(p)

# p[30]
p[30]

# barplot
barplot(p,names.arg=x,xlab="Total Selections",ylab="Probability")
```

b.  Using `which`, `sum`, and various logical conditions, determine the
    following probabilities.

-   P(X=3); about 10.7%
-   P(X \> 40); about 9.1%
-   P(X is between 10 and 20); 10 and 20 are included; about 16.4%
-   P(X is an odd number); about 53%; hint, use a `seq` command
-   P( X is at most 10 or between 15-20 or at least 45 ); about 59%;
    hint, you'll want three logical conditions separated by "or"s

```{r Q2b}
#P(X=3)
sum(p[which(x == 3)])

#P(X > 40)
sum(p[which(x > 40)])

#P(X is between 10 and 20 inclusive)
sum(p[which(x >= 10 & x <= 20)])

#P(X is an odd number)
sum(p[which(x %% 2 == 1)])

#P( X is at most 10 or  15-20 or at least 45 )
sum(p[which(x <= 10 | (x >= 15 & x <= 20) | x >= 45)])
```

------------------------------------------------------------------------

## Question 3

Half-way into your workday, you get stuck reading some data into R. You
email your analytics team hoping they can clarify something about the
data. Waiting to hear back from them can feel like forever! Let `X` be
the number of minutes it takes to receive a response (let's round to the
nearest minute so it's all integer values). The possible values of `x`
are 0, 1, 2, 3, ... (no real upper limit, so we'll take that to be
"infinity").

$$Shape = x^2 e^{-x/20} \qquad x=0, 1, 2, 3, \ldots$$

This shape is not a valid PMF since the sum of all the "shape" values
does not equal 1. It's not possible to do the standard shape conversion
trick (`p <- shape/sum(shape)`) because there are an infinite number of
possible values of `x`. However, it *is* possible to convert the shape
equation into a PMF by multiplying it by the reciprocal of a well-chosen
number!

$$P(X=x) = \frac{1}{???}\left( x^2 e^{-x/20}  \right) \qquad x=0, 1, 2, 3, \ldots$$

a)  Use Wolfram-Alpha to find the value of `???` in the equation above.
    Wolfram will give the "exact" answer which involves a lot of weird
    powers of "e", and then it'll say it's $\approx$ equal to some
    value. Feel free to use that approximate value for the remainder of
    the problem! Provide a screenshot of your work in Wolfram (paste
    into Word after knitting).

**Response:** 16000 (SCREENSHOT HERE)

b)  Although the values of `x` technically go up to infinity, define `x`
    to be the integer sequence 1 to 300 and `p` to be the vector of
    corresponding probabilities *using your equation* for $P(X=x)$
    (after you've solved for `???` in (a); no `p <- shape/sum(shape)`
    here!)

b1) Show `sum(p)` is "basically 1" (I was getting 0.9999615), implying
that `x` and `p` provide a "reasonable enough" PMF for the quantity we
wish to study since we've captured "most" of the probabilities (even
though the possible values of `x` go to infinity).

b2) Provide a bar chart of the PMF. Label the x-axis "Minutes until
response" and the y-axis "Probability".

Remember that to take `e` and raise it to a power in R, you need to
write it as `exp()`, with the power going inside the parentheses.

```{r Q3b12}
x = 1:300
p = (1 / 16000) * ((x^2) * (exp(-x/20)))
sum(p)
barplot(p, names.arg=x, xlab="Minutes until response", ylab="Probability")
```

c)  Use your PMF vectors `x` and `p` to find the probability that you
    have to wait at least 2 hours for a response. Sanity check: a little
    over 6.3%.

```{r Q3c}
sum(p[which(x >= (2 * 60))])
```

------------------------------------------------------------------------

## Question 4

Let's practice identifying when to use various notorious discrete
probability distributions. Choose between uniform, geometric, binomial,
negative binomial, Poisson, and zero-Truncated Poisson. Do not choose
negative binomial when the geometric distribution could be used (the
geometric is simpler, more specific case of one flavor of the negative
binomial).

a.  What would be the go-to distribution to model the number of times a
    mobile food-delivering robot collides with a person or other robot
    over the course of a week?

**Response:** Poisson distribution.

b.  What would be the go-to distribution to model the number of members
    of the UT's Sports Analytics Society (92 members) who own both a dog
    and a cat?

**Response:** Binomial distribution.

c.  What would be the go-to distribution to model the number of Rick and
    Morty references that Brian Steven makes in a STAT 201 or BAS 320
    lecture?

**Response:** Zero-Truncated Poisson distribution.

d.  What would be the go-to distribution to model the total number of
    times a person has donated to UT (among only those people who
    previously donated to UT)?

**Response:** Poisson distribution.

e.  What would be the go-to distribution to model the minute of the hour
    (1-60) of someone's birth?

**Response:** Uniform distribution.

f.  What would be the go-to distribution to model the number of credit
    and debit cards someone has in their wallet (not everyone carries
    these cards)?

**Response:** Zero-Truncated Poission distribution.

g.  What would be the go-to distribution to model the number of
    deliveries you receive from amazon.com before you receive a delivery
    that contains a defective product?

**Response:** Geometric distribution.

h.  What would be the go-to distribution to model the number of the
    orange M & Ms (Go Vols!) in a typical 1.69oz pack (56 candies
    total)?

**Response:** Binomial distribution.

i.  Each day, a dog has a probability `p` of trying to escape the
    backyard. Other days (with probability `1-p`), the dog stays put and
    does not try to escape. On each attempt at escape, the dog learns a
    bit more as to how to get out. Imagine that on the 8th escape
    attempt, the dog will finally be able to get out. What is the go-to
    distribution to model the total number of days that the dog stays
    put in the yard (with no attempt at escape) before the dog finally
    gets out?

**Response:** Geometric distribution.

------------------------------------------------------------------------

## Question 5:

Let's practice using the notorious discrete probability distributions in
R.

a.  Let's model the number of times that shoppers of Best Buy make a
    purchase in-store or online in a December with a Poisson
    distribution with an average of 0.9. Determine the value of `lambda`
    for this Poisson (hint: this is an easy one), then determine

-   the probability someone makes exactly 2 purchases in December
-   the probability someone makes the average number of purchases (0.9)
    in December
-   the probability that makes 4 or more purchases in December

```{r Q5a}
#2 purchases
dpois(2, 0.9)

#Average number of purchases
# this answer would be 0 because someone can't make a decimal number of purchases.
0

#4 or more purchases
1 - ppois(3, 0.9)
```

b.  Most (but not all) delivery attempts by the food-delivering robots
    on campus are completed in a timely manner. Let's define a
    "successful delivery" to be one that takes at most 15 minutes.
    Imagine that the probability a robot is able to make a successful
    delivery is 0.88, and that the result of each delivery is
    independent of one another. If the robot is scheduled to make 60
    deliveries in day, find the probability that:

-   the robot has a perfect record for the day (all 60 deliveries are 15
    minutes or less) ; about 0.05%
-   the robot has between 48 and 56 (inclusive) successful deliveries;
    about 91.6%
-   the robot has 50 or fewer successful deliveries; about 17.8%

```{r Q5b}
#Exactly 60
dbinom(60, 60, 0.88)

#Between 48-56 inclusive
sum(dbinom(48:56, 60, 0.88))

#50 or fewer
pbinom(50, 60, 0.88)
```

c.  You are asked to give out flyers about the Sports Analytics Club to
    people in the student union. Oddly enough, most people are not
    interested, with each person independently having a 93.5% chance of
    refusing a flyer. After quite some time, you end up with only a
    single flyer left to give away.

The geometric distribution describes the number of people we have to ask
*before* someone finally accepts that last flyer. For the following
questions, figure out what it's asking about the number of "failures"
(refusals of the flyer) before the first "success" (accepting the
flyer), then find the requested probabilities.

-   the probability that you will give away the last flyer to the 6th
    person you ask (about 4.6%).\
-   the probability that you'll give away your last flier to the 4th,
    5th, 6th, 7th, 8th, or 9th person you ask (about 27.1%)
-   the probability you'll have to ask *more than 20* people before you
    run into the person who takes your last flyer (about 24.4%)

```{r Q5c}
#Last flyer to the 6th person you ask
dgeom(6, 1 - 0.935)

#Last flyer to person 4-9 inclusive (2 different ways to do it)
sum(dgeom(4:9, c(1 - 0.935,1 - 0.935,1 - 0.935,1 - 0.935,1 - 0.935, 1 - 0.935)))
pgeom(9, 1- 0.935) - pgeom(3, 1 - 0.935)

#More than 20 people
sum(1 - pgeom(20, 1- .935))
```

d.  Selling in-game currency for real-world money in online games like
    World of Warcraft and Final Fantasy XIV is referred to as RMT, or
    "real money trading". It's a nice way to earn some extra cash, but
    because it is expressly forbidden in the terms of service, doing so
    risks your account getting banned from the game. Imagine each RMT
    sale has a 20% chance of being detected. An account gets banned once
    the 3rd RMT sale has been detected. Find the probability that:

-   You are able to make at most 10 RMT sales before your account gets
    banned, i.e. the 3rd detection happens on the 10th sale or before
    (about 32.2%)

-   You are able to make at least 30 RMT sales before your account gets
    banned, i.e. the 3rd detection happens on the 30th sale or later
    (about 5.2%)

Remember that the negative binomial is counting up the number of
*failures* before a certain number of successes. Treat a "failure" as an
undetected RMT sale and a "success" as a detected RMT sale (these
definitions feel awkward from *our* perspective since we want the
transactions to be undetected, it does make sense from the game owner's
perspective!).

Translate these questions into how many failures are required to
describe the event, then use the `nbinom` suite of functions.

```{r Q5d}
#Make at most 10 RMT sales before account gets banned. 
pnbinom(7, 3, .20)

#Make at least 30 RMT sales
# 1 - at most 29 = at least 30, 29 - 3 = 26
1- pnbinom(26, 3, .20)
```

------------------------------------------------------------------------

## Question 6 (bonus)

As part of protest, batters are refusing to swing at pitches. The result
of the game will thus come down to pitcher skill. Imagine that each
pitch has a 65% chance of being a "strike" and a 35% chance of being a
"ball". A batter has a "strike out" if 3 strikes occur before 4 balls. A
batter has a "walk" if 4 balls occur before 3 strikes. You've seen a
similar problem before back in an activity done in class for Unit 1, and
in an illustration in a Unit 2 Monte Carlo simulation, but this problem
takes the analysis much further.

a)  What's the name of the "notorious" distribution that counts up the
    number of failures that occur before a target number of successes?

**Response:** Negative binomial distribution.

b)  Let a ball be considered a "failure" and a strike a "success". A
    strikeout occurs when there are 3 strikes (successes) *before* 4
    balls (failures). In other words, a strikeout occurs when the number
    of "failures" before the 3rd success is at most 3. Use the
    appropriate `d` or `p` function in R to calculate the probability of
    a strikeout (remember back in Unit 1 how tedious it was the
    enumerate all possible ways a strikeout could occur?). It turns out
    to be close to 88%.

```{r Q6b}
pnbinom(1, 1, .65)
```

c)  Imagine a similar game called "buskerball" is invented where a
    batter has a "strike out" if 4 strikes occur before 6 balls. A
    batter has a "walk" if 6 balls occur before 4 strikes. If the
    probability of throwing a strike is 0.65, what's the probability of
    a strikeout? It turns out to be close to 95%.

```{r Q6c}
pnbinom(2, 1, .65)
```

d)  Back to baseball. Although I haven't seen this as an official sports
    analytics interview question, I wouldn't be surprised if it is one
    for some team. In protest, no one is swinging their bat during a
    game. An "inning" will end with no runs scored as long as 3
    strikeouts occur before 4 walks. Use one of the notorious
    distributions to find the probability an inning has no runs scored
    (about 99.8%). Then, find the probability that the game ends without
    any runs being scored (9 innings in a row result in no runs scored;
    about 97.9%).

```{r Q6d}

```

e)  Unfortunately, the pitcher who has the 65% chance of throwing a
    strike is injured and had to be replaced by a pitcher that only has
    a 50% chance of throwing a strike.

e1) Find the PMF (`x` and `p`) of the number of points scored in an
inning, i.e., - The probability an inning ends with 0 points scored (3
strikeouts occur before 4 walks) - The probability an inning ends with 1
point scores (the 3rd strikeout occurs after exactly 4 walks) - The
probability an inning ends with 2 point scores (the 3rd strikeout occurs
after exactly 5 walks) - etc.\
- Let `x` be `0:15` (so you go up to 15 points scored) and `p` be the
corresponding probabilities. Print out the contents of your `p` vector
once you've defined them, then make a barplot to show the PMF. Have the
x-label be "Points Scored in Inning" and y-label be "Probability"

e2) Simulate the PMF of the total number of runs scored over the course
of a game (9 innings) by sampling (with replacement) 9 values from the
PMF for runs scored in an inning. - Run 100,000 trials, storing the
total number of points scored in the game in a vector named
`total.points` - Include a barplot of the results, i.e.,
`barplot(table(total.points))`. Have the x-label be "Points Scored in
Game" and y-label be "Probability" - What is the probability that at
least 5 points are scored during these 9 innings (about 11%)?

```{r Q6e}

```
